2040 greenhouse gas reduction targets and energy transitions in line with the EU Green Deal

### 2025-04-24 report_paperCharts.Rmd
update limVRE setup and general cleanup

# all emission reductions were calculated relative to 1990 emissions with LULUCF and international transport (4814 Mt CO2e), except for 2030 emission reduction calculated relative to 1990 emissions with LULUCF and intra-EU aviation only (4713 Mt CO2e).

@@

g[["GHG emi"]] <- ggplot(plotData, aes(x=period,y=value)) +
    geom_hline(yintercept=0, color = "black", linewidth=2, alpha = 0.5) +
    geom_vline(xintercept=2020, color = "black", linewidth=1.5, alpha = 0.5,linetype="dashed") +
    geom_ribbon(aes(ymax = ribbon.max, ymin = ribbon.min), alpha = 0.5, fill = "#545454") + 
    geom_line(data=plotData %>% filter(scenario == mainScen), linewidth=1.5) +
    #geom_line(data=plotData, color= "#545454", linewidth=0.2, alpha=0.2, aes(group=scenario)) +
    geom_line(data=h,linewidth=0.7,linetype="dashed") +
    geom_line(data=linearProjection,linewidth=0.8,linetype="dashed") +
    #geom_segment(data=plotData %>% filter(scenario == mainScen, period == 2040), aes(x=2040, xend=2040, y=0, yend=value), linetype="dashed", color="black") +
    geom_segment(data=linearProjection %>% filter(period == 2040), aes(x=2040, xend=2040, y=0, yend=value), linetype="dashed", color="black") +
    geom_point(data=plotData %>% filter(scenario == mainScen, period == 2040),size=3,shape=21,fill="white",stroke=1.5) +
    geom_point(data=linearProjection %>% filter(period == 2040),size=3,shape=21,fill="white",stroke=1.5) +
    geom_label_repel(data=data.frame(period=c(2042.5),value=mean(linearProjection %>% filter(period %in% c(2040,2045)) %>% pull(value))),
                     label = "linear\nreduction",
                     nudge_x = 0,
                     nudge_y = 2500,
                     segment.square = TRUE,
                     point.padding = 0,
                     hjust = 0.5,
                     direction="y",
                     size = geomText.size) + 
    theme_doubleColumn() +
    theme(axis.title.y=element_blank()) +
    expand_limits(y=emi1990) +
    scale_y_continuous(breaks = c(0,1000,2000,3000,4000,5000)) +
    scale_x_continuous(breaks = seq(1990,2050,10)) +
    labs(title = expression(paste("(a) Total GHG emissions (Mt ", CO[2],"eq/yr)"))) + 
    #theme(plot.title = element_text(hjust = -0.2)) +
    theme(plot.title = element_text(margin=margin(b=35)))

@@

g[["target"]] <- ggplot(data=plotData,aes(x=period,y=reduction,fill=bar)) +
  geom_bar(data=plotData %>% filter(scenario == mainScen), stat='identity', show.legend = FALSE, alpha = 0.7, color="black") +
  geom_errorbar(aes(ymin=min, ymax=max), width=1, colour="black", alpha=0.9, linewidth=1.3) +
  geom_hline(yintercept=0, color = "black", linewidth=2, alpha = 0.5) +
  theme_doubleColumn() +
  scale_fill_manual(values = c("emi" = "black", "2040" = "#5a788b")) + 
  guides(y = "none") +
  geom_segment(data=plotData %>% filter(scenario == mainScen), aes(x=x, y=y, xend=xend, yend=yend), arrow = arrow(length=unit(.5, 'cm')), na.rm=TRUE) +
  geom_text(data=plotData %>% filter(scenario == mainScen), aes(label = label, y=labelY, x=labelX, color=factor(period)), vjust=-0.2, hjust=0.5, na.rm=TRUE, size = geomText.size * 0.8, angle = 90) +
  scale_color_manual(values = c("2030" = "black", "2035" = "#777777","2040" = "black", "2045" = "#777777","2050" = "black"), breaks=c("2030", "2035","2040", "2045","2050")) +
  expand_limits(y=minValue/emi1990) +
  annotate("text", x=2052, y=1, label= "1990 emissions", vjust=-1, hjust=1, size = geomText.size) +
  geom_hline(yintercept=1, color = "black", linetype="dashed", linewidth=1.5, alpha = 0.5) +
  coord_cartesian(clip = 'off') +
  scale_y_continuous(breaks = c(0,1000/emi1990,2000/emi1990,3000/emi1990,4000/emi1990,5000/emi1990)) + 
  theme(legend.position = "none") +
  labs(title = "(b) Emission reductions (%)") + 
  theme(plot.title = element_text(margin=margin(b=35)))

g[["GHG Emission reductions"]] <- gridExtra::arrangeGrob(
  g[["GHG emi"]], #+ theme(plot.margin=margin(10,10,30,10)),
  g[["target"]], # + theme(plot.margin=margin(10,10,10,10)),
  layout_matrix = rbind(c(1, 1, 1, 1, 2, 2, 2),
                        c(1, 1, 1, 1, 2, 2, 2))
   )
# 
#  ggsave(paste0(outChartsFolder,"/svg/Figure 1. Emission results panel - (a) GHG and (b) reduction.svg"), g[["GHG Emission reductions"]] , device="svg", width = 12, height = 5, dpi=100, units = "in")
#  ggsave(paste0(outChartsFolder,"/png/Figure 1. Emission results panel - (a) GHG and (b) reduction.png"), g[["GHG Emission reductions"]] , device="png", width = 12, height = 5, dpi=300, units = "in")

@@

vars <- c(
  "Energy Supply - Electricity"               = "Emi|GHG|Gross|Energy|Supply|Electricity",
  "Energy Supply - Non-electric"              = "Emi|GHG|Gross|Energy|Supply|Non-electric",
  "Energy Demand - Transport"                 = "Emi|GHG|Energy|Demand|Transport",
  "Energy Demand - International Bunkers"     = "Emi|CO2|Energy|Demand|Transport|International Bunkers",
  "Energy Demand - Buildings"                 = "Emi|GHG|Energy|Demand|Buildings",
  "Energy Demand - Industry and Agriculture"  = "Emi|GHG|Gross|Energy|Demand|Industry",
  "Energy Demand - Carbon Dioxide Removal"    = "Emi|GHG|Energy|Demand|CDR",
  "Industrial Processes"                      = "Emi|GHG|Industrial Processes",
  "Agriculture"                               = "Emi|GHG|Agriculture",
  "Land-Use Change"                           = "Emi|GHG|Land-Use Change|LULUCF national accounting", #"Emi|GHG|Land-Use Change",
  "Waste"                                     = "Emi|GHG|Waste",
  "BECCS"                                     = "Emi|CO2|CDR|BECCS",
  "Industry CCS"                              = "Emi|CO2|CDR|Industry CCS|Synthetic Fuels",
  "DACCS"                                     = "Emi|CO2|CDR|DACCS",
  "Enhanced Weathering"                       = "Emi|CO2|CDR|EW")

plotColor <- c(
  "Energy Supply - Electricity"               = "#ffb200",
  "Energy Supply - Non-electric"              = "#0000cc",
  "Energy Demand - Transport"                 = "#a1dd70",
  "Energy Demand - International Bunkers"     = "#ffc0cb",
  "Energy Demand - Buildings"                 = "#5edfff",
  "Energy Demand - Industry and Agriculture"  = "#a5a5a5",
  "Energy Demand - Carbon Dioxide Removal"    = "#999959",
  "Industrial Processes"                      = "#5c5c5c",
  "Agriculture"                               = "#357933",
  "Land-Use Change"                           = "#6ec16b",
  "Waste"                                     = "#A52A2A",
  "BECCS"                                     = "#377EB8",
  "Industry CCS"                              = "#5ed5b0",
  "DACCS"                                     = "#CC6666",
  "Enhanced Weathering"                       = "#5b0d8f")

@@

g[["GHG Emissions per sector"]] <- ggplot(plotData,aes(x=period,y=value)) +
  geom_area(aes(fill=variable), alpha = 0.8) +
  geom_line(data=totalData, linewidth = 1, color="black") +
  geom_vline(xintercept=2020, color = "black", linetype="dashed", linewidth=1, alpha = 0.5) +
  geom_hline(yintercept=0, color = "black", linewidth=2, alpha = 0.5) +
  theme_doubleColumn() +
  #ylab("Final Energy Demand - EJ/yr") + 
  #ggtitle("(a) per sector and carrier") +
  scale_fill_manual(values = setNames(plotColor[names(vars)],vars), labels = setNames(names(vars), vars)) +
  #theme(legend.margin=margin(l = -1, unit='cm')) +
  scale_x_continuous(breaks=c(2010, 2010, 2020,2030,2040,2050)) + 
  theme(axis.title.x=element_blank()) +
  theme(axis.title.y=element_blank()) +
  theme(legend.title = element_blank()) +
  theme(#legend.text = element_text(size = theme.double.text.size*0.7),
        #legend.key.height = unit(0.7, "cm"),
        #legend.key.width = unit(0.7,"cm"),
        legend.spacing.x = unit(0.2, 'cm')) + 
  #  theme(plot.title = element_text(size = 1.2*theme.double.text.size)) +
  theme(legend.position="right", legend.direction = "vertical", legend.box = "vertical") +
  labs(title = "(c) GHG Emissions per sector in the reference scenario (Mt CO2/yr)") + 
  theme(legend.margin=margin(t=30)) +
  theme(plot.margin = margin(t = 15, unit = "pt"))

@@

vars <- list(
  "Industrial\nProcesses" = "Emi|GHG|Industrial Processes",
  "Industry" = "Emi|GHG|Gross|Energy|Demand|Industry",
  "Transportation" ="Emi|GHG|Energy|Demand|Transport",
  "Bunkers" = "Emi|CO2|Energy|Demand|Transport|International Bunkers",
  "Buildings" ="Emi|GHG|Energy|Demand|Buildings",
  "Energy\nSupply"="Emi|CO2|Gross|Energy|Supply"
)

plotColor <- c(
  "Industrial\nProcesses" = "#5c5c5c",
  "Industry" = "#5f6769",
  "Transportation" = "#a1dd70",
  "Bunkers" = "#ffc0cb",
  "Buildings" = "#5edfff",
  "Energy\nSupply"= "#ffb200"
)

@@

histVars <- list(
  "Industrial\nProcesses" = "Emi|GHG|Industrial Processes",
  "Industry" = "Emi|GHG|Industry",
  "Transportation" ="Emi|GHG|Energy|Demand|Transport",
  "Bunkers" = "Emi|GHG|Energy|Demand|Transport|International Bunkers",
  "Buildings" ="Emi|GHG|Energy|Demand|Buildings",
  "Energy\nSupply"="Emi|CO2|Gross|Energy|Supply"
)

@@

  geom_text(data=plotData %>% filter(period == 2040), aes(x=xmid , y = 2045, label = pct1990), vjust=0.5, hjust=-0.1, na.rm=TRUE, size = geomText.size, angle = 90) +
  scale_fill_manual(values = c(setNames(plotColor[names(vars)],vars)), labels = setNames(names(vars), vars)) + 
  scale_alpha_manual(values = c(0.2, 0.2, 0.3, 1)) +
  scale_color_manual(values = c(setNames(plotColor[names(vars)],vars)), labels = setNames(names(vars), vars)) +
  theme_doubleColumn() +
  theme(legend.position = "none") +
  scale_y_continuous(breaks=c(2010,2020,2030,2040), labels=c("1990","avrg. 2018-2022","2030","2040"), minor_breaks = NULL, limits=c(1998,2047)) +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank()) +
  scale_x_continuous(breaks = NULL) +
  coord_cartesian(clip="off") +
  theme(plot.margin = margin(t = 15, r = 0, b = 100, l = 0, unit = "pt")) +
  #theme(plot.margin = margin(t = 30, r = 0, b = 100, l = 0, unit = "pt")) +
  labs(title = "(d) Sectoral breakdown of residual net CO2 emissions from energy and industrial processes") + 
  theme(plot.title = element_text(margin=margin(t=0,r=0,b=55,l=0), hjust=1))

@@

#label <- c(

#)

@@

g[["Electricity Generation"]] <- 
  ggplot(plotData,aes(x=period,y=percentage)) +
  geom_area(aes(fill=variable),alpha=0.4) +
  geom_line(data=plotDataRenewables, linetype = "dashed", linewidth = 1.5, color="#666666") +
  geom_bar(data=plotDataBars,aes(fill=variable),stat='identity', show.legend = FALSE, alpha = 0.7, color="white", width = 3, linewidth = 1.5) +
  geom_text(data=plotDataBars %>% filter(percentage >= 0.02),aes(label = label, y=cumPercentage-percentage/2, x = xPos), vjust=0.5, hjust=0, na.rm=TRUE, color = "black", size=geomText.size*0.8) +
  #geom_text(data=plotData2030[plotData2030$percentage >= 0.02,],aes(label = label, y=cumPercentage-percentage/2, x = xPos), vjust=0.5, hjust=0, na.rm=TRUE, color = "black", size=geomText.size*0.8) +
  #geom_text(data=plotData2040[plotData2040$percentage >= 0.02,],aes(label = paste0(round(percentage,2)*100,"%", ifelse(inv_vars[variable] %in% c("Solar PV","Wind Onshore","Wind Offshore"),paste0(" - ", inv_vars[variable]), "")), y=cumPercentage-percentage/2, x = xPos), vjust=0.5, hjust=0, na.rm=TRUE, color = "black", size=geomText.size) +
  theme_doubleColumn()  +
  scale_y_continuous(minor_breaks = seq(0.2, 0.8, 0.2), breaks=c(0,1),labels = scales::percent) +
  scale_x_continuous(breaks=c(2020,2030,2040,2050)) +
  scale_fill_manual(values = setNames(color[names(vars)],vars), labels = setNames(names(vars), vars)) +
  guides(fill=guide_legend(ncol=2,override.aes = list(alpha=1))) + #, byrow = TRUE
  annotate(geom = "text", x = 2015, y = plotDataRenewables[plotDataRenewables$period == 2015,]$percentage/2, label = "renewables\nand nuclear", color = "#666666",angle = 90, hjust=0.5, vjust=-0.5, size=geomText.size, lineheight = 0.8) +
  annotate(geom = "text", x = 2015, y = (1-plotDataRenewables[plotDataRenewables$period == 2020,]$percentage)/2+plotDataRenewables[plotDataRenewables$period == 2015,]$percentage, label = "fossil", color = "#666666",angle = 90, hjust=0.5, vjust=-2, size=geomText.size) +
  annotation_custom(grob = linesGrob(gp=gpar(col="#666666",lwd = 3,lty="dashed")), xmin = 2011, xmax = 2014.8, ymin = plotDataRenewables[plotDataRenewables$period == 2015,]$percentage, ymax = plotDataRenewables[plotDataRenewables$period == 2015,]$percentage) +
  #annotate(geom = "text", label = "100%", x = 2020, y = 1, hjust=1.2, color = "#666666", size=geomText.size) +
  #annotate(geom = "text", label = "0%", x = 2020, y = 0, hjust=1.2, color = "#666666", size=geomText.size) +
  coord_cartesian(clip = 'off') +
  #theme(plot.margin = margin(t = 0, r = 0, b = 0, l = 20, unit = "pt")) +
  labs(title = "(a) Electricity Generation (%)") #+ 
  #theme(plot.title = element_text(hjust = -0.1))

# ggsave(paste0(outChartsFolder,"/svg/Figure 2. Electricity panel - (a) generation.svg"), g[["Electricity Generation"]] , device="svg", width = 6, height = 10, dpi=100, units = "in")
# ggsave(paste0(outChartsFolder,"/png/Figure 2. Electricity panel - (a) generation.png"), g[["Electricity Generation"]] , device="png", width = 6, height = 10, dpi=300, units = "in")

# g[["Electricity Generation - wide"]] <- 
#   ggplot(plotData,aes(x=period,y=percentage)) +
#   geom_area(aes(fill=variable),alpha=0.4) +
#   geom_line(data=plotDataRenewables, linetype = "dashed", linewidth = 1.5, color="#666666") +
#   geom_bar(data=plotData2040,aes(fill=variable),stat='identity', show.legend = FALSE, alpha = 0.7, color="white", width = 3, linewidth = 1.5) +
#   geom_text(data=plotData2040[plotData2040$percentage >= 0.02,],aes(label = paste0(round(percentage,2)*100,"%", ifelse(inv_vars[variable] %in% c("Solar PV","Wind Onshore","Wind Offshore"),paste0(" - ", inv_vars[variable]), "")), y=cumPercentage-percentage/2, x = xPos), vjust=0.5, hjust=0, na.rm=TRUE, color = "black", size=geomText.size) +
#   theme_doubleColumn()  +
#   scale_y_continuous(minor_breaks = seq(0.2, 0.8, 0.2), breaks=c(0,1),labels = scales::percent) +
#   scale_x_continuous(breaks=c(2020,2040,2050)) +
#   scale_fill_manual(values = setNames(color[names(vars)],vars), labels = setNames(names(vars), vars)) +
#   theme(legend.position="right") +
#   theme(legend.margin=margin(b = -2, unit='cm'))

#ggsave(paste0(outFolder_conference,"/Electricity Generation - wide.svg"), g[["Electricity Generation - wide"]] , device="svg", width = 12, height = 5, dpi=100, units = "in")
#ggsave(paste0(outFolder_conference,"/Electricity Generation - wide.png"), g[["Electricity Generation - wide"]] , device="png", width = 12, height = 5, dpi=300, units = "in")

@@

g[["Solar_Photovoltaics_capacity"]] <- ggplot(plotData,aes(x=period,y=value)) +
    geom_hline(yintercept=0, color = "black", linewidth=2, alpha = 0.5) +
    geom_vline(xintercept=2020, color = "black", linewidth=1.5, alpha = 0.5,linetype="dashed") +
    geom_segment(aes(x=2040, y=0, xend=2040, yend=yendValue), linetype="dashed", color="black") +
    geom_ribbon(aes(ymax = ribbon.max, ymin = ribbon.min), alpha = 0.5, fill = "#545454") + 
    geom_line(data=plotData %>% filter(scenario == mainScen),linewidth=1.5) +
    geom_line(data=h,linewidth=1,linetype="dashed") +
    geom_point(data=plotData %>% filter(scenario == mainScen, period %in% c(2030,2040)),size=3,shape=21,fill="white",stroke=1.5) +
    geom_point(data=target, size=3, shape=4, stroke = 1.5) +
    geom_label_repel(data=target, aes(label = paste0(round(value),"GW - ", target)),box.padding = 0.1,
                  #segment.curvature = -0.1,
                  segment.square = TRUE,
                  point.padding = 0.1,
                  nudge_x = c(4,1.5),
                  #nudge_y = -0.02,
                  #force = 0.5,
                  hjust = 0,
                  direction="y", size = geomText.size * 0.8, xlim = c(2020, 2070)) +
  geom_label_repel(data=plotData %>% filter(scenario == mainScen, period == 2030), aes(label = paste0(round(value),"GW")),box.padding = 0.1,
                  point.padding = 1,
                  nudge_x = -2,
                  nudge_y = 500,
                  hjust = 1, direction="y",
                  size = geomText.size) +
  geom_label_repel(data=plotData %>% filter(scenario == mainScen, period == 2040), aes(label = paste0(round(value),"GW")),box.padding = 0.1,
                  point.padding = 1,
                  nudge_x = -2,
                  nudge_y =500,
                  #hjust = 0, 
                  direction="y",
                  size = geomText.size * 0.8) +
    theme_doubleColumn() + 
    coord_cartesian(clip = 'off') +
    expand_limits(y = 3000) +
    labs(title = "(b) Solar Photovoltaics capacity (GW)") #+ 
    #theme(plot.title = element_text(hjust = -0.4))

# ggsave(paste0(outChartsFolder,"/svg/Figure 2. Electricity panel - (b) solar.svg"), g[["Solar_Photovoltaics_capacity"]] , device="svg", width = 6, height = 4, dpi=100, units = "in")
# ggsave(paste0(outChartsFolder,"/png/Figure 2. Electricity panel - (b) solar.png"), g[["Solar_Photovoltaics_capacity"]] , device="png", width = 6, height = 4, dpi=300, units = "in")

@@

# with range
plotData <- df %>% 
  filter(region == "EU27", period %in% c(2020,2030,2040,2050), tgt2030 %in% tgt2030Scen, variable %in% unlist(vars), tgt2050 != "Npi") %>% 
  #mutate(variable = factor(variable, levels=vars))  %>%
  mutate(type = ifelse(grepl("Imports",variable),"Imports","Exports"),
         carrier = carrier[as.character(variable)]) %>%
  filter(carrier != "Biomass") %>%
  mutate(value = ifelse(type == "Exports", -value,value)) %>%
  mutate(carrier = factor(carrier, levels=c("Oil","Gas","Coal","Biomass"))) %>%
  group_by(model,scenario,region,unit,period,carrier) %>%
  summarize(value=sum(value),.groups = "keep") %>%
  mutate(type="EU-27 Net-trade") %>%
  group_by(model,region,carrier,period) %>% 
  mutate(ribbon.max=max(value),
         ribbon.min=min(value))

@@

}

>>>>

# Emission reductions relative to 1990 emissions include LULUCF and intra-EU aviation only (4713 Mt CO2e), except for 2045 and 2050 emission reductions which are calculated relative to 1990 emissions including LULUCF and international transport (4814 Mt CO2e).

@@

g[["GHG emi"]] <- ggplot(plotData, aes(x=period,y=value)) +
    geom_hline(yintercept=0, color = "black", linewidth=2, alpha = 0.5) +
    geom_vline(xintercept=2020, color = "black", linewidth=1.5, alpha = 0.5,linetype="dashed") +
    geom_ribbon(aes(ymax = ribbon.max, ymin = ribbon.min), alpha = 0.5, fill = "#545454") + 
    geom_line(data=plotData %>% filter(scenario == mainScen), linewidth=1.5) +
    geom_line(data=h,linewidth=0.7,linetype="dashed") +
    geom_line(data=linearProjection,linewidth=0.8,linetype="dashed") +
    geom_segment(data=linearProjection %>% filter(period == 2040), aes(x=2040, xend=2040, y=0, yend=value), linetype="dashed", color="black") +
    geom_point(data=plotData %>% filter(scenario == mainScen, period == 2040),size=3,shape=21,fill="white",stroke=1.5) +
    geom_point(data=linearProjection %>% filter(period == 2040),size=3,shape=21,fill="white",stroke=1.5) +
    geom_label_repel(data=data.frame(period=c(2042.5),value=mean(linearProjection %>% filter(period %in% c(2040,2045)) %>% pull(value))),
                     label = "linear\nreduction",
                     nudge_x = 0,
                     nudge_y = 2500,
                     segment.square = TRUE,
                     point.padding = 0,
                     hjust = 0.5,
                     direction="y",
                     size = geomText.size) + 
    expand_limits(y=emi1990) +
    scale_y_continuous(breaks = c(0,1000,2000,3000,4000,5000)) +
    scale_x_continuous(breaks = seq(1990,2050,10)) +
    labs(title = expression(paste("(a) Total GHG emissions (Mt ", CO[2],"eq/yr)"))) + 
    theme_doubleColumn() +
    theme(axis.title.y=element_blank(),
          plot.title = element_text(margin=margin(b=35)))

@@

g[["target"]] <- ggplot(data=plotData,aes(x=period,y=reduction,fill=bar)) +
  geom_bar(data=plotData %>% filter(scenario == mainScen), stat='identity', show.legend = FALSE, alpha = 0.7, color="black") +
  geom_errorbar(aes(ymin=min, ymax=max), width=1, colour="black", alpha=0.9, linewidth=1.3) +
  geom_hline(yintercept=0, color = "black", linewidth=2, alpha = 0.5) +
  scale_fill_manual(values = c("emi" = "black", "2040" = "#5a788b")) + 
  guides(y = "none") +
  geom_segment(data=plotData %>% filter(scenario == mainScen), aes(x=x, y=y, xend=xend, yend=yend), arrow = arrow(length=unit(.5, 'cm')), na.rm=TRUE) +
  geom_text(data=plotData %>% filter(scenario == mainScen), aes(label = label, y=labelY, x=labelX, color=factor(period)), vjust=-0.2, hjust=0.5, na.rm=TRUE, size = geomText.size * 0.8, angle = 90) +
  scale_color_manual(values = c("2030" = "black", "2035" = "#777777","2040" = "black", "2045" = "#777777","2050" = "black"), breaks=c("2030", "2035","2040", "2045","2050")) +
  expand_limits(y=minValue/emi1990) +
  annotate("text", x=2052, y=1, label= "1990 emissions", vjust=-1, hjust=1, size = geomText.size) +
  geom_hline(yintercept=1, color = "black", linetype="dashed", linewidth=1.5, alpha = 0.5) +
  coord_cartesian(clip = 'off') +
  scale_y_continuous(breaks = c(0,1000/emi1990,2000/emi1990,3000/emi1990,4000/emi1990,5000/emi1990)) + 
  labs(title = "(b) Emission reductions (%)") + 
  theme_doubleColumn() +
  theme(legend.position = "none",
        plot.title = element_text(margin=margin(b=35)))

@@

vars <- c(
  "Energy Supply - Electricity"               = "Emi|GHG|Gross|Energy|Supply|Electricity",
  "Energy Supply - Non-electric"              = "Emi|GHG|Gross|Energy|Supply|Non-electric",
  "Energy Demand - Transport"                 = "Emi|GHG|Energy|Demand|Transport",
  "Energy Demand - International Bunkers"     = "Emi|CO2|Energy|Demand|Transport|International Bunkers",
  "Energy Demand - Buildings"                 = "Emi|GHG|Energy|Demand|Buildings",
  "Energy Demand - Industry and Agriculture"  = "Emi|GHG|Gross|Energy|Demand|Industry",
  "Energy Demand - CDR"                       = "Emi|GHG|Energy|Demand|CDR",
  "Industrial Processes"                      = "Emi|GHG|Industrial Processes",
  "Agriculture"                               = "Emi|GHG|Agriculture",
  "Land-Use Change"                           = "Emi|GHG|Land-Use Change|LULUCF national accounting", #"Emi|GHG|Land-Use Change",
  "Waste"                                     = "Emi|GHG|Waste",
  "BECCS"                                     = "Emi|CO2|CDR|BECCS",
  "Industry CCS"                              = "Emi|CO2|CDR|Industry CCS|Synthetic Fuels",
  "DACCS"                                     = "Emi|CO2|CDR|DACCS",
  "Enhanced Weathering"                       = "Emi|CO2|CDR|EW")

plotColor <- c(
  "Energy Supply - Electricity"               = "#ffb200",
  "Energy Supply - Non-electric"              = "#0000cc",
  "Energy Demand - Transport"                 = "#a1dd70",
  "Energy Demand - International Bunkers"     = "#ffc0cb",
  "Energy Demand - Buildings"                 = "#5edfff",
  "Energy Demand - Industry and Agriculture"  = "#a5a5a5",
  "Energy Demand - CDR"                       = "#999959",
  "Industrial Processes"                      = "#5c5c5c",
  "Agriculture"                               = "#357933",
  "Land-Use Change"                           = "#6ec16b",
  "Waste"                                     = "#A52A2A",
  "BECCS"                                     = "#377EB8",
  "Industry CCS"                              = "#5ed5b0",
  "DACCS"                                     = "#CC6666",
  "Enhanced Weathering"                       = "#5b0d8f")

@@

g[["GHG Emissions per sector"]] <- ggplot(plotData,aes(x=period,y=value)) +
  geom_area(aes(fill=variable), alpha = 0.8) +
  geom_line(data=totalData, linewidth = 1, color="black") +
  geom_vline(xintercept=2020, color = "black", linetype="dashed", linewidth=1, alpha = 0.5) +
  geom_hline(yintercept=0, color = "black", linewidth=2, alpha = 0.5) +
  scale_fill_manual(values = setNames(plotColor[names(vars)],vars), labels = setNames(names(vars), vars)) +
  scale_x_continuous(breaks=c(2010, 2010, 2020,2030,2040,2050)) + 
  labs(title = "(c) GHG Emissions per sector in the reference scenario (Mt CO2/yr)") + 
  theme_doubleColumn() +
  theme(axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        legend.title = element_blank(),
        legend.spacing.x = unit(0.2, 'cm'),
        legend.position="right", 
        legend.direction = "vertical", 
        legend.box = "vertical",
        legend.margin=margin(t=30),
        plot.margin = margin(t = 15, unit = "pt"))

@@

vars <- list(
  "Industrial\nProcesses" = "Emi|GHG|Industrial Processes",
  "Industry" = "Emi|GHG|Gross|Energy|Demand|Industry",
  "Transport" ="Emi|GHG|Energy|Demand|Transport",
  "Bunkers" = "Emi|CO2|Energy|Demand|Transport|International Bunkers",
  "Buildings" ="Emi|GHG|Energy|Demand|Buildings",
  "Energy\nSupply"="Emi|CO2|Gross|Energy|Supply"
)

plotColor <- c(
  "Industrial\nProcesses" = "#5c5c5c",
  "Industry" = "#5f6769",
  "Transport" = "#a1dd70",
  "Bunkers" = "#ffc0cb",
  "Buildings" = "#5edfff",
  "Energy\nSupply"= "#ffb200"
)

@@

histVars <- list(
  "Industrial\nProcesses" = "Emi|GHG|Industrial Processes",
  "Industry" = "Emi|GHG|Industry",
  "Transport" ="Emi|GHG|Energy|Demand|Transport",
  "Bunkers" = "Emi|GHG|Energy|Demand|Transport|International Bunkers",
  "Buildings" ="Emi|GHG|Energy|Demand|Buildings",
  "Energy\nSupply"="Emi|CO2|Gross|Energy|Supply"
)

@@

  geom_text(data=plotData %>% filter(period == 2040), aes(x=xmid , y = 2045, label = pct1990), vjust=0.5, hjust=-0.1, na.rm=TRUE, size = geomText.size, angle = 90) +
  scale_fill_manual(values = c(setNames(plotColor[names(vars)],vars)), labels = setNames(names(vars), vars)) + 
  scale_alpha_manual(values = c(0.2, 0.2, 0.3, 1)) +
  scale_color_manual(values = c(setNames(plotColor[names(vars)],vars)), labels = setNames(names(vars), vars)) +
  scale_y_continuous(breaks=c(2010,2020,2030,2040), labels=c("1990","avrg. 2018-2022","2030","2040"), minor_breaks = NULL, limits=c(1998,2047)) +
  scale_x_continuous(breaks = NULL) +
  coord_cartesian(clip="off") +
  #theme(plot.margin = margin(t = 30, r = 0, b = 100, l = 0, unit = "pt")) +
  labs(title = "(d) Sectoral breakdown of residual net CO2 emissions from energy and industrial processes") + 
  theme_doubleColumn() +
  theme(legend.position = "none",
        axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
        plot.margin = margin(t = 15, r = 0, b = 100, l = 0, unit = "pt"),
        plot.title = element_text(margin=margin(t=0,r=0,b=55,l=0), hjust=1))

@@

@@

g[["Electricity Generation"]] <- 
  ggplot(plotData,aes(x=period,y=percentage)) +
  geom_area(aes(fill=variable),alpha=0.4) +
  geom_line(data=plotDataRenewables, linetype = "dashed", linewidth = 1.5, color="#666666") +
  geom_bar(data=plotDataBars,aes(fill=variable),stat='identity', show.legend = FALSE, alpha = 0.7, color="white", width = 3, linewidth = 1.5) +
  geom_text(data=plotDataBars %>% filter(percentage >= 0.02),aes(label = label, y=cumPercentage-percentage/2, x = xPos), vjust=0.5, hjust=0, na.rm=TRUE, color = "black", size=geomText.size*0.8) +
  scale_y_continuous(minor_breaks = seq(0.2, 0.8, 0.2), breaks=c(0,1),labels = scales::percent) +
  scale_x_continuous(breaks=c(2020,2030,2040,2050)) +
  scale_fill_manual(values = setNames(color[names(vars)],vars), labels = setNames(names(vars), vars)) +
  guides(fill=guide_legend(ncol=2,override.aes = list(alpha=1))) + #, byrow = TRUE
  annotate(geom = "text", x = 2015, y = plotDataRenewables[plotDataRenewables$period == 2015,]$percentage/2, label = "renewables\nand nuclear", color = "#666666",angle = 90, hjust=0.5, vjust=-0.5, size=geomText.size, lineheight = 0.8) +
  annotate(geom = "text", x = 2015, y = (1-plotDataRenewables[plotDataRenewables$period == 2020,]$percentage)/2+plotDataRenewables[plotDataRenewables$period == 2015,]$percentage, label = "fossil", color = "#666666",angle = 90, hjust=0.5, vjust=-2, size=geomText.size) +
  annotation_custom(grob = linesGrob(gp=gpar(col="#666666",lwd = 3,lty="dashed")), xmin = 2011, xmax = 2014.8, ymin = plotDataRenewables[plotDataRenewables$period == 2015,]$percentage, ymax = plotDataRenewables[plotDataRenewables$period == 2015,]$percentage) +
  coord_cartesian(clip = 'off') +
  labs(title = "(a) Electricity Generation (%)") +
  theme_doubleColumn()

@@

g[["Solar_Photovoltaics_capacity"]] <- ggplot(plotData,aes(x=period,y=value)) +
    geom_hline(yintercept=0, color = "black", linewidth=2, alpha = 0.5) +
    geom_vline(xintercept=2020, color = "black", linewidth=1.5, alpha = 0.5,linetype="dashed") +
    geom_segment(aes(x=2040, y=0, xend=2040, yend=yendValue), linetype="dashed", color="black") +
    geom_ribbon(aes(ymax = ribbon.max, ymin = ribbon.min), alpha = 0.5, fill = "#545454") + 
    geom_line(data=plotData %>% filter(scenario == mainScen),linewidth=1.5) +
    geom_line(data=h,linewidth=1,linetype="dashed") +
    geom_point(data=plotData %>% filter(scenario == mainScen, period %in% c(2030,2040)),size=3,shape=21,fill="white",stroke=1.5) +
    geom_point(data=target, size=3, shape=4, stroke = 1.5) +
    geom_label_repel(data=target, aes(label = paste0(round(value),"GW - ", target)),box.padding = 0.1,
                  #segment.curvature = -0.1,
                  segment.square = TRUE,
                  point.padding = 0.1,
                  nudge_x = c(4,1.5),
                  #nudge_y = -0.02,
                  #force = 0.5,
                  hjust = 0,
                  direction="y", size = geomText.size * 0.8, xlim = c(2020, 2070)) +
  geom_label_repel(data=plotData %>% filter(scenario == mainScen, period == 2030), aes(label = paste0(round(value),"GW")),box.padding = 0.1,
                  point.padding = 1,
                  nudge_x = -2,
                  nudge_y = 500,
                  hjust = 1, direction="y",
                  size = geomText.size * 0.8) +
  geom_label_repel(data=plotData %>% filter(scenario == mainScen, period == 2040), aes(label = paste0(round(value),"GW")),box.padding = 0.1,
                  point.padding = 1,
                  nudge_x = -2,
                  nudge_y =500,
                  #hjust = 0, 
                  direction="y",
                  size = geomText.size * 0.8) +
    coord_cartesian(clip = 'off') +
    expand_limits(y = 3000) +
    labs(title = "(b) Solar Photovoltaics capacity (GW)") + 
    theme_doubleColumn()

@@

# with range
plotData <- df %>% 
  filter(region == "EU27", period %in% c(2020,2030,2040,2050), tgt2030 %in% tgt2030Scen, variable %in% unlist(vars), tgt2050 != "Npi") %>% 
  mutate(type = ifelse(grepl("Imports",variable),"Imports","Exports"),
         carrier = carrier[as.character(variable)]) %>%
  filter(carrier != "Biomass") %>%
  mutate(value = ifelse(type == "Exports", -value,value)) %>%
  mutate(carrier = factor(carrier, levels=c("Oil","Gas","Coal","Biomass"))) %>%
  group_by(model,scenario,region,unit,period,carrier) %>%
  summarize(value=sum(value),.groups = "keep") %>%
  mutate(type="EU-27 Net-trade") %>%
  group_by(model,region,carrier,period) %>% 
  mutate(ribbon.max=max(value),
         ribbon.min=min(value))

@@
